A simplified dynamic bioenergetic model for coral-\emph{Symbiodinium} symbioses

### 2016-09-20 JTB_ms/JTB_ms.Rmd
update Fig2

bibliography: library.bib
output: 
  rticles::elsevier_article:
    fig_caption: yes
---
```{r setup, include=FALSE}
#knitr::opts_knit$set(root.dir = normalizePath(".."))
```

@@

To analyze qualitative behavior, we ran the model to steady state across gradients of external irradiance and DIN, and plotted the specific growth rate (Fig. 2A) and symbiont to host biomass ratio (Fig. 2B). Results are consistent with observations in corals: specific growth rates

![Steady state values of (A) specific growth ($Cmol \cdot Cmol^{-1} \cdot d^{-1}$) and (B) the symbiont to host biomass ratio ($CmolS \cdot CmolH^{-1}$) across gradients of external irradiance and dissolved inorganic nitrogen. Simulations for each combination of light and nutrients (21 points along each axis) were run for 100 days with a time step of 1 day. Negative steady state growth rates, and corresponding S:H ratios, were set to zero.](../img/Fig2.png)

>>>>

bibliography: library.bib
output: 
  rticles::elsevier_article:
    fig_caption: yes
---
```{r setup, include=FALSE}
library(reshape2); library(dplyr)
#knitr::opts_knit$set(root.dir = normalizePath(".."))
```

@@

To analyze qualitative behavior, we ran the model to steady state across gradients of external irradiance and DIN, and plotted the specific growth rate (Fig. 2A) and symbiont to host biomass ratio (Fig. 2B). Results are consistent with observations in corals: specific growth rates are low at very low levels of irradiance and DIN, as these scenarios limit the availability of carbon and nitrogen, respectively, to build biomass. As these resources increase, specific growth increases, reaching its highest levels above 2 µM DIN and ~15 mol photons m^-2^ d^-1^. Peak specific growth rates are around ~0.1 d^-1^, which is consistent with values reported in the literature (@Muscatine:1981jy).

![Steady state values of (A) specific growth (Cmol Cmol^-1^ d^-1^) and (B) the symbiont to host biomass ratio (CmolS CmolH^-1^) across gradients of external irradiance and dissolved inorganic nitrogen. Simulations for each combination of light and nutrients (21 points along each axis) were run for 100 days with a time step of 1 day. Negative steady state growth rates, and corresponding S:H ratios, were set to zero.](../img/Fig2.png)

### 2016-09-20 JTB_ms/JTB_ms.tex
update Fig2

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{longtable}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A simplified dynamic bioenergetic model for coral-Symbiodinium symbioses},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
\usepackage{tabularx}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header
\usepackage{tabularx}

@@

\begin{figure}[htbp]
\centering
\includegraphics{JTB_ms_files/figure-latex/Fig1-1.pdf}
\caption{Graphical representation of coral-algal symbiosis model.}
\end{figure}

@@

where \(j_{HGm}\) is the maximum specific growth rate. Finally, the rate
of change in coral biomass is calculated as the specific growth rate
minum the specific biomass turnover rate (i.e., maintenance costs),
according to:

\begin{equation} {dH \over dt} = j_{HG} - j_{HT}^0 \end{equation}

@@

This second component of symbiont biomass loss can represent both
photodamage and/or symbiont expulsion (i.e., bleaching), both of which
happen in response to higher levels or ROS production. The constant 5 is
included to increase biomass loss in response to ROS. (Note that
recycling of symbiont biomass turnover (\(rNS\) and \(rCS\)) only occurs
based on the basal maintenance related turnover (i.e., \(j_{ST}^0\)),
and not the bleaching-related biomass loss, as this loss represents
biomass being damaged or being expelled from the holobiont). Finally,
the balance equation representing the change in symbiont biomass over
time can be specified as the growth rate minus the turnover rate:

\begin{equation} {dS \over dt} = j_{SG} - j_{ST} \end{equation}

@@

@@

\hypertarget{refs}{}
\hypertarget{ref-Costanza:2014ex}{}
Costanza, R, R de Groot, and P Sutton. 2014. ``Changes in the global
value of ecosystem services.'' \emph{Global Environmental \ldots{}} 26:
152--58.
doi:\href{https://doi.org/10.1016/j.gloenvcha.2014.04.002}{10.1016/j.gloenvcha.2014.04.002}.

@@

\hypertarget{ref-Glynn:2001p7571}{}
Glynn, Peter W, Juan L Maté, Andrew C Baker, and MO Calderón. 2001.
``Coral bleaching and mortality in Panama and Ecuador during the
1997-1998 El Niño-Southern Oscillation event: Spatial/temporal patterns
and comparisons with the 1982-1983 event.'' \emph{Bulletin of Marine
Science} 69 (1): 79--109.
\url{http://www.scopus.com/record/display.url?fedsrfIntegrator=MEKPAPERS-SCOCIT\&origin=fedsrf\&view=basic\&eid=2-s2.0-0034761826}.

@@

\hypertarget{ref-Jokiel:1977p7353}{}
Jokiel, Paul L, and SL Coles. 1977. ``Effects of temperature on the
mortality and growth of Hawaiian reef corals.'' \emph{Marine Biology} 43
(3): 201--8.
\url{http://www.scopus.com/record/display.url?fedsrfIntegrator=MEKPAPERS-SCOCIT\&origin=fedsrf\&view=basic\&eid=2-s2.0-0000237488}.

@@

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{longtable}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A simplified dynamic bioenergetic model for coral-Symbiodinium symbioses},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

\begin{figure}[htbp]
\centering
\includegraphics{../img/model.png}
\caption{Graphical representation of coral-algal symbiosis model.}
\end{figure}

@@

@@

This second component of symbiont biomass loss can represent both
photodamage and/or symbiont expulsion (i.e., bleaching), both of which
happen in response to higher levels or ROS production. The constant 5 is
included to increase biomass loss in response to ROS. (Note that
recycling of symbiont biomass turnover (\(rNS\) and \(rCS\)) only occurs
based on the basal maintenance related turnover (i.e., \(j_{ST}^0\)),
and not the bleaching-related biomass loss, as this loss represents
biomass being damaged or being expelled from the holobiont).

\subsection{Model state equations}\label{model-state-equations}

Finally, the balance equations representing the specific growth rates of
symbiont and host biomass over time can be expressed as:

\begin{equation} {dS \over Sdt} = j_{SG} - j_{ST} \end{equation}

\begin{equation} {dH \over Hdt} = j_{HG} - j_{HT}^0 \end{equation}

\subsection{Numerical analysis}\label{numerical-analysis}

A time-stepping Euler method was used to solve the state equations since
the production and rejection fluxes of the SUs are imiplicitly defined.
Specifically, the rejection fluxes of carbon and nitrogen from the
symbiont and host biomass SUs act as reciprocal input fluxes to the
other SU. In addition, the rejection flux of excitation energy from the
photosynthesis SU acts to reduce its own production flux (i.e.,
photoinhibition), and hence a discretized time-stepping procedure was
necessary. A vector of time values was created for each simulation run,
along which dynamic environmental forcing functions (irradiance, DIN,
and prey abundance) can be designed. These vectors, along with initial
values of symbiont and host biomass, then serve as input to the
time-stepping function, which solves for the current system state using
values of the previous system state where necessary. A default time step
of 0.1 days was used for all simulations, which were performed using R
code that is available in the data repository accompanying this article:
\url{github.com/jrcunning/Rcoral}.

@@

To analyze qualitative behavior, we ran the model to steady state across
gradients of external irradiance and DIN, and plotted the specific
growth rate (Fig. 2A) and symbiont to host biomass ratio (Fig. 2B).
Results are consistent with observations in corals: specific growth
rates are low at very low levels of irradiance and DIN, as these
scenarios limit the availability of carbon and nitrogen, respectively,
to build biomass. As these resources increase, specific growth
increases, reaching its highest levels above 2 µM DIN and
\textasciitilde{}15 mol photons m\textsuperscript{-2}
d\textsuperscript{-1}. Peak specific growth rates are around
\textasciitilde{}0.1 d\textsuperscript{-1}, which is consistent with
values reported in the literature (Muscatine, R McCloskey, and E Marian
(1981)).

\begin{figure}[htbp]
\centering
\includegraphics{../img/Fig2.png}
\caption{Steady state values of (A) specific growth (Cmol
Cmol\textsuperscript{-1} d\textsuperscript{-1}) and (B) the symbiont to
host biomass ratio (CmolS CmolH\textsuperscript{-1}) across gradients of
external irradiance and dissolved inorganic nitrogen. Simulations for
each combination of light and nutrients (21 points along each axis) were
run for 100 days with a time step of 1 day. Negative steady state growth
rates, and corresponding S:H ratios, were set to zero.}
\end{figure}

@@

\hypertarget{refs}{}
\hypertarget{ref-Costanza:2014ex}{}
Costanza, R, R de Groot, and P Sutton. 2014. ``Changes in the global
value of ecosystem services.'' \emph{Global Environmental Change} 26:
152--58.
doi:\href{https://doi.org/10.1016/j.gloenvcha.2014.04.002}{10.1016/j.gloenvcha.2014.04.002}.

@@

\hypertarget{ref-Glynn:2001p7571}{}
Glynn, Peter W, Juan L Maté, Andrew C Baker, and MO Calderón. 2001.
``Coral bleaching and mortality in Panama and Ecuador during the
1997-1998 El Niño-Southern Oscillation event: Spatial/temporal patterns
and comparisons with the 1982-1983 event.'' \emph{Bulletin of Marine
Science} 69 (1): 79--109.

@@

\hypertarget{ref-Jokiel:1977p7353}{}
Jokiel, Paul L, and SL Coles. 1977. ``Effects of temperature on the
mortality and growth of Hawaiian reef corals.'' \emph{Marine Biology} 43
(3): 201--8.

@@

\hypertarget{ref-Muscatine:1981jy}{}
Muscatine, Leonard, L R McCloskey, and R E Marian. 1981. ``Estimating
the daily contribution of carbon from zooxanthellae to coral animal
respiration.'' \emph{Limnology and Oceanography} 26 (4): 601--11.
doi:\href{https://doi.org/10.4319/lo.1981.26.4.0601}{10.4319/lo.1981.26.4.0601}.
